### 2019-02-22 scripts/00_introduction.Rmd
FIX title on 00_introduction.Rmd

---
title: "Introduction: a pipeline to analyse time-course gene expression data"
output: html_document
---

# Abstract

The phenotypic diversity of cells is governed by a complex equilibrium between
their genetic identity and their environmental interactions, and understanding
the dynamics of gene expression is a fundamental question of biology. However,
analysing time-course transcriptomic data raises challenging statistical and
computational questions, requiring the development of novel methods and
software.

# Introduction

- TODO why is time-course gene expression analysis important for biology. Why
  are novel workflows necessary.

Time-course gene expression analysis workflows include steps used more broadly
for gene expression data analasys such as quality control (QC), exploratory
data analysis (EDA), pre-processing, normalization, and visualization. Yet, a
typical gene expression data analysis workflow is not entirely adequate for
analysing time-course data. In particular, differential expression analysis
needs to be adapted to time-course data, and often needs to be followed by a
clustering step to group genes sharing similar patterns of expression.

Here, we propose an integrated workflow for downstream analysis of time-course
data, with the following main steps: (1) quality control and normalization;
(2) differential expression analysis of time-course data; (3) clustering; (4)
downstream analysis of clusters with GO term and pathway enrichment.

XXX Add the packages used in the workflow

# Analysis of the dynamical response of mouse lung tissue to influenza

## Overview

This workflow is illustrated using data from a micro-array time-course
experiment, exposing mice to three different strains of influenza, and
collecting lung tissue during 14 time-points after infection (0, 3, 6, 9, 12,
18, 24, 30, 36, 48, 60 hours, then 3, 5, and 7 days later). The three strains
of influenza used in the study are (1) a low pathogenicity seasonal H1N1
influenza virus (A/Kawasaki/UTK4/2009 [H1N1]), a mildly pathogenic virus from
the 2009 pandemic season (A/California/04/2009 [H1N1]), and a highly
pathogenic H5N1 avian influenza virus (A/Vietnam/1203/2004 [H5N1]. Mice were
injected with 10^5 PFU of each virus. An adidtional 42 mice were injected with
a lower dose of the Vietnam avian influenza virus (10^3 PFU).

**XXX Add summary of the key findings of the the paper**