### 2019-03-06 scripts/05_conclusion.Rmd

---
title: "Conclusion: A pipeline to analyse time-course gene expression data"
output: rmarkdown::html_vignette
---

# Author contributions

NV and EP wrote the workflow.

# Competing interests
The authors declare that they have no competing interests.

# Grant information

This research was funded in part by a Department of Energy (DOE) grant
(DE-SC0014081); by the Gordon and Betty Moore Foundation (Grant GBMF3834) and
the Alfred P. Sloan Foundation (Grant 2013-10-27) to the University of
California, Berkeley [N.V.]; by a ENS-CFM Data Science Chair [E.P.].

*I confirm that the funders had no role in study design, data collection and
analysis, decision to publish, or preparation of the manuscript.*

# Acknowledgments

The authors thank Karthik Ram and the Ropensci community for valuable
feedback.