### 2019-03-17 scripts/02_differential_expression_analysis.Rmd
ENH Updating the rest of the functions to use the object

---
title: "Differential expression analysis: A pipeline to analyse time-course gene expression data"
output: rmarkdown::html_vignette
---

```{r echo=FALSE, results="hide"}
# Reloading necessary data and preliminary results
library(devtools)
library(limma)
library(splines)
library(stats)
library(moanin)
```

```{r echo=FALSE}
# Set options
df = 6

# Preprocessing & filtering
is_count = FALSE
take_log = FALSE    
filter_expression = FALSE

# Differential expression analysis
timecourse_contrasts = c("M-K", "M-C", "M-VL", "M-VH")

# Clustering
filter_genes = TRUE
n_genes_to_keep = 5000
n_clusters = 20
percentage_genes_to_label = 0.5
```

```{r echo=FALSE, results="hide"}
# Set data files and options
data("shoemaker2015")
data = shoemaker2015$data
meta = shoemaker2015$meta

# Right now, the package doesn't contain the normalized data
```

```{r, echo=FALSE, results="hide"}
# XXX FIXME
data = read.table(".results/quantile_normalized.txt") 
source("utils.R")
```
```{r}
splines_model = moanin::create_splines_model(meta, degrees_of_freedom=6)
```

## Differential expression analysis of time-course data

The next step is typically to run a differential expression analysis.

### Weekly differential expression analysis

`moanin` provides a simple interface to perform a time by time differential
expression analysis. Under the hood, it simply calls `limma` on the set of
contrasts provided. Here, we show an example between the control mouse and the
mouse infected with the high dose of the influenza strain A/Vietnam/1203/04
(H5N1), but the function works with any form contrasts.

```{r weekly_differential_analysis}
meta$WeeklyGroup = make.names(meta$Group:as.factor(meta$Time))

# Define contrast	
contrasts = c("M.0-VL.0",
	      "M.3-VL.3",
	      "M.6-VL.6",
	      "M.9-VL.9",
	      "M.12-VL.12",
	      "M.18-VL.18",
	      "M.24-VL.24",
	      "M.30-VL.30",
	      "M.36-VL.36",
	      "M.48-VL.48",
	      "M.60-VL.60",
	      "M.72-VL.72",
	      "M.120-VL.120",
	      "M.168-VL.168")
weekly_de_analysis =  moanin::weekly_differential_expression_analysis(
     data, splines_model, contrasts,
     use_voom_weights=FALSE)
```

```{r echo=FALSE, results="hide"}
# Hide those results
# Run the DE analysis on the rest of the influenza strain.
contrasts = c("M.0-K.0",
	      "M.3-K.3",
	      "M.6-K.6",
	      "M.9-K.9",
	      "M.12-K.12",
	      "M.18-K.18",
	      "M.24-K.24",
	      "M.30-K.30",
	      "M.36-K.36",
	      "M.48-K.48",
	      "M.60-K.60",
	      "M.72-K.72",
	      "M.120-K.120",
	      "M.168-K.168")
K_weekly_de_analysis =  moanin::weekly_differential_expression_analysis(
     data, splines_model, contrasts,
     use_voom_weights=FALSE)
contrasts = c("M.0-C.0",
	      "M.3-C.3",
	      "M.6-C.6",
	      "M.9-C.9",
	      "M.12-C.12",
	      "M.18-C.18",
	      "M.24-C.24",
	      "M.30-C.30",
	      "M.36-K.36",
	      "M.48-C.48",
	      "M.60-C.60",
	      "M.72-C.72",
	      "M.120-C.120",
	      "M.168-C.168")
C_weekly_de_analysis =  moanin::weekly_differential_expression_analysis(
     data, splines_model, contrasts,
     use_voom_weights=FALSE)
contrasts = c("M.0-VH.0",
	      "M.3-VH.3",
	      "M.6-VH.6",
	      "M.9-VH.9",
	      "M.12-VH.12",
	      "M.18-VH.18",
	      "M.24-VH.24",
	      "M.30-VH.30",
	      "M.36-VH.36",
	      "M.48-VH.48",
	      "M.60-VH.60",
	      "M.72-VH.72",
	      "M.120-VH.120",
	      "M.168-VH.168")
VH_weekly_de_analysis =  moanin::weekly_differential_expression_analysis(
     data, splines_model, contrasts,
     use_voom_weights=FALSE)

```

Let's look at the distribution of genes found differentially expressed per
week between control and each of the influenza strains.

```{r fig.width=8, fig.height=8, echo=FALSE}
par(mfrow=c(2, 2))

time = c(0, 3, 6, 9, 12, 18, 24, 30, 36, 48, 60, 72, 120, 168)

timely_barplot = function(de_analysis, labels, qval_threshold=0.05, title=""){
    qval_colnames = colnames(de_analysis)[
	grepl("qval", colnames(de_analysis))]
    number_de_genes_per_time = colSums(de_analysis[, qval_colnames] < 0.05)
    names(number_de_genes_per_time) = labels
    barplot(number_de_genes_per_time, col="black", ylim=c(0, 20000),
	    xlab="Time", ylab="Number of DE genes", main=title)

}

timely_barplot(K_weekly_de_analysis, time, title="Kawasaki")
timely_barplot(C_weekly_de_analysis, time, title="California")
timely_barplot(weekly_de_analysis, time, title="Vietnam low-dose")
timely_barplot(VH_weekly_de_analysis, time, title="Vietnam high-dose")
```

The distribution of number of genes found differentially expressed by
considering each time-point independantly highlights the challenges of such
approach. Timepoint 6H and 18H of the Kawasaki strain have fewer numbers of
genes found differentially expressed than timepoints 3H and 9H: this is likely
due to biological or technical variances for specific genes at specific
timepoints.

As a summary, classic differential expression methods are appropriate for
unordered treatments, but fail to use the temporal structure of the data.

### Time-course differential expression analysis between two groups

To leverage this temporal structure, Storey et al [CITE FIXME] proposes to
model time-course micro-array with splines functions, and to use a log-ratio
likelihood test to detect differentially expressed genes. `moanin` extends
this idea by providing functions to perform time-course differential
expression analysis using `limma` contrasts.

```{r de_analysis}
# Define contrast
contrasts = limma::makeContrasts(
	contrasts=timecourse_contrasts,
	levels=levels(meta$Group))

# The function takes the data (data.frame or named matrix), the meta data
# (data.frame containing a timepoint and group column, the first corresponding
# to the time-course information, the latter corresponding to the
# treatment).
pvalues = moanin::timecourse_differential_expression_analysis(
    data, splines_model, contrasts)
qvalues = apply(pvalues, 2, stats::p.adjust)

log_fold_change_epicon = moanin::estimate_log_fold_change(
    data, splines_model, contrasts,  method="epicon")

log_fold_change_max = moanin::estimate_log_fold_change(
    data, splines_model, contrasts, method="max")
```

Let's explore a bit the number of genes found differentially expressed.

```{r echo=FALSE, fig.width=6}
num_de = colSums(qvalues < 0.05)
names(num_de) = c(
    "Kawasaki", "California", "Vietnam low-dose", "Vietname high-dose")
barplot(num_de, col="black", ylab="Number of DE genes")

```

And now a volcano plot for quick visualization

```{r fig.height=8, fig.width=8}

pvalue = qvalues[, "M-K"]
lfc_epicon = log_fold_change_epicon[, "M-K"]

plot(lfc_epicon, -log10(pvalue), pch=20, main="Volcano plot",
     xlim=c(-2.5, 2))

# Add colored points: red if padj<0.05, orange of log2FC>1, green if both)
points(lfc_epicon[pvalue < 0.01], -log10(pvalue[pvalue < 0.01]), pch=20,
       col="red")
points(lfc_epicon[abs(lfc_epicon) > 1],
       -log10(pvalue[abs(lfc_epicon) > 1]), pch=20, col="orange")
points(lfc_epicon[(pvalue < 0.01) & (abs(lfc_epicon) > 1)],
       -log10(pvalue[(pvalue < 0.01) & (abs(lfc_epicon) > 1)]), pch=20, col="green")

```

```{r echo=FALSE, results="hide"}
# Rename stuff for saving
outname = ".results/pvalues.txt"
colnames(pvalues) = unlist(
    lapply(colnames(pvalues), function(x){paste0(x, "-pval")}))

colnames(qvalues) = unlist(
    lapply(colnames(qvalues), function(x){paste0(x, "-qval")}))

colnames(log_fold_change_epicon) = unlist(
    lapply(colnames(log_fold_change_epicon),
    function(x){paste0(x, "-lfc_epicon")}))

colnames(log_fold_change_max) = unlist(
    lapply(colnames(log_fold_change_max),
    function(x){paste0(x, "-lfc_max")}))

de_analysis = merge(pvalues, qvalues, all=TRUE, by="row.names")
row.names(de_analysis) = row.names(pvalues)
de_analysis = subset(de_analysis, select=-c(Row.names))

de_analysis = merge(de_analysis, log_fold_change_max, all=TRUE, by="row.names")
row.names(de_analysis) = row.names(pvalues)
de_analysis = subset(de_analysis, select=-c(Row.names))

de_analysis = merge(de_analysis, log_fold_change_epicon, all=TRUE,
		    by="row.names")
row.names(de_analysis) = row.names(pvalues)
de_analysis = subset(de_analysis, select=-c(Row.names))

write.table(de_analysis, outname, sep="\t")
```