### 2019-03-17 scripts/01_quality_control_normalization.Rmd
MAINT rename Time -> Timepoint

---
title: "Normalization: A pipeline to analyse time-course gene expression data"
output: rmarkdown::html_vignette
---

```{r echo=FALSE, results="hide", message=FALSE}
# Loading dependencies
library(devtools)
library(limma)
library(splines)
library(stats)
library(moanin)
library(ggplot2)
library(stats)

source("utils.R")

# library(moanin)
```

```{r echo=FALSE}
# Set options
df = 6

# Preprocessing & filtering
is_count = FALSE
take_log = FALSE    
filter_expression = FALSE

# Differential expression analysis
timecourse_contrasts = c("C-K", "C-M")

# Clustering
filter_genes = TRUE
n_genes_to_keep = 5000
n_clusters = 20
percentage_genes_to_label = 0.5
```

```{r echo=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, error=FALSE, message=FALSE,
  echo=TRUE,
  duplicate.label="allow"
)

```

## Pre-processing and normalization

The micro-array data were obtained from NCBI Gene Expression Omnibus (GEO),
with accession number GSE63786. The dataset has a total of 209 samples, with
45018 probes (1 sample was removed by the authors prior to uploading the
dataset on GEo).

Note that in the following, we assume that the user has access to a data
folder `data/shoemaker2015/rawdata`, containing the microarray data.

```{r}
data_dir = "data/shoemaker2015/rawdata"
```

The normalization strategy we apply here relies on `limma` and is as follows:

  - First, normalize probe intensity using background information using the
    "norm-exponential" strategy.
  - Then normalize across micro-arrrays using quantile normalization.

```{r reading_targets, results="hide", fig.width=8, fig.height=4}
# Read in the data using limma.
targets = limma::readTargets()
RG = limma::read.maimages(
  targets, source="agilent",
  path=data_dir,
  other.columns="gIsWellAboveBG",
  green.only=TRUE)

# Preform background correction using norm-exponential strategy and normalize
# across samples using quantile normalization.
RG = limma::backgroundCorrect(RG, method="normexp")
RG.q = limma::normalizeBetweenArrays(RG, method="quantile")
```

Plotting the smoothed empirical densities of the probe values before
normalization shows considerable variations betweens arrays. Once the data has
been normalized across array using quantile normalization, the smoothed
empirical densities of each sample perfectly overlap.

```{r plotting_densities, results="hide", fig.width=8, fig.height=4}
# Plot the densities of the probes before and after normalization.
par(mfrow=c(1, 2))
limma::plotDensities(RG, col="black", legend=FALSE)
limma::plotDensities(RG.q, col="black", legend=FALSE)
```

```{r echo=FALSE, results="hide"}
# Get the number for the text. Do not display results or computation
control_probes = RG.q$genes[, "ControlType"] == 1
is_expressed = rowSums(RG.q$other$"gIsWellAboveBG" > 0) >= 4
n_control_probes = sum(control_probes)
n_lowly_expressed = sum(!is_expressed)
```

We then clean up the names of each gene and each sample to match the original
data. Finally, we further filter out probes based on the following criteria:
(1) removing the `r n_control_probes` control probes; (2) removing `r
n_lowly_expressed` low-expressed genes.

```{r clean_up_names}

sample_names = sapply(strsplit(colnames(RG.q), "_"), .subset2,  1)
colnames(RG.q) = sample_names
row.names(RG.q) = make.names(RG$genes$SystematicName, unique=TRUE)
idx_to_keep = which(
  !(RG$genes$Systematic %in% c("DarkCorner", "GE_BrightCorner",
			       "NegativeControl")))

data = RG.q[idx_to_keep,]

# Filter out control probes.
control_probes = data$genes[, "ControlType"] == 1
is_expressed = rowSums(data$other$"gIsWellAboveBG" > 0) >= 4

data = data[!control_probes & is_expressed, ]$E
```

 XXX The original paper says the data is "averaged over unique probes IDs."

```{r echo=FALSE, results="hide"}
# Do not display. setting variables for text
n_samples = dim(data)[2]
n_probes = dim(data)[1]
```

After filtering, we are left with `r n_probes` probes of interest in `r
n_samples`.

```{r}
# Now load in the metadata
# meta = read.delim("data/shoemaker2015/mice_meta.tsv", sep="\t")
data(shoemaker2015)
meta = shoemaker2015$meta
write.table(data, ".results/quantile_normalized.txt")
```

## Exploratory analysis and quality control

Plotting the smoothed empirical densities of probe values for each sample and
estimating the variations across samples before and after normalization is a
first quality control. Typically, two other quality control and exploratory
analysis steps are also performed before and after normalization: (1) low
dimensionality embedding of the samples; (2) correlation plots between each
samples. In both cases, we expect a strong biological signal, while replicate
samples should be strongly clustered or correlated with one another.

Before performing any additional exploratory analysis, let us only keep highly
variable genes.

```{r}
variance_cutoff = 0.5
variance_per_genes = apply(data, 1, mad)
min_variance = stats::quantile(variance_per_genes, c(variance_cutoff))
data = data[variance_per_genes > min_variance,]
```

```{r echo=FALSE, fig.width=8, fig.height=4}
graphics::hist(variance_per_genes, breaks=100, col="black", xlab="Variance",
	       ylab="# probes")
graphics::abline(v=min_variance, col="#AB0000", lw=2)
```

Let us first perform the PCA analysis. Here, we perform a PCA of rank 3 of the
centered and scaled gene expression data.

```{r pca, fig.width=8, fig.height=8}
# Reorder genes on condition, time, and replicate

pca_data = stats::prcomp(t(data), rank=3, center=TRUE, scale=TRUE) 
percent_var = round(100 * attr(pca_data, "percentVar"))
```

We then plot the two first components, and color each sample by (1) its
condition; (2) its sampling time. We use different markers for each replicate.
We also plot the second and third components in the second row.

```{r pca_plots, fig.width=8, fig.height=8}
# FIXME Add legend
graphics::par(mfrow=c(2, 2))
graphics::plot(
    pca_data$x[, "PC2"], pca_data$x[, "PC1"],
    col=ann_colors$Group[meta$Group],
    pch=ann_markers$Replicate[as.factor(meta$Replicate)],
    xlab="PC2", ylab="PC1")
graphics::plot(
    pca_data$x[, "PC2"], pca_data$x[, "PC1"],
    col=ann_colors$Timepoint[as.factor(meta$Timepoint)],
    pch=ann_markers$Replicate[as.factor(meta$Replicate)],
    xlab="PC2", ylab="PC1")

graphics::plot(
    pca_data$x[, "PC2"], pca_data$x[, "PC3"],
    col=ann_colors$Group[meta$Group],
    pch=ann_markers$Replicate[as.factor(meta$Replicate)],
    xlab="PC2", ylab="PC3")
graphics::plot(
    pca_data$x[, "PC2"], pca_data$x[, "PC3"],
    col=ann_colors$Timepoint[as.factor(meta$Timepoint)],
    pch=ann_markers$Replicate[as.factor(meta$Replicate)],
    xlab="PC2", ylab="PC3")
```

Then, we plot the pearson correlation across each samples. We order the
samples by their Group (treatment) and Timepoint (the time of sampling).
Additionally, in this example, we order each treatment by strength of the
pathogenicity of the treatment: Control, Kawasaki, California, low dose of
Vietnam, then high dose of Vietnam.

```{r  correlation_plot, fig.width=6, fig.height=6}
library(NMF)

# Reorder genes on condition, time, and replicate
meta$Group = factor(meta$Group, levels(meta$Group)[c(3, 1, 2, 5, 4)])
ord = order(
  meta$Group,
  meta$Timepoint,
  meta$Replicate)
data = data[, ord]
meta = meta[ord, ]

data_corr = stats::cor(data, method="pearson")

meta$Timepoint = as.factor(meta$Timepoint)
NMF::aheatmap(
    data_corr,
    Colv=NA, Rowv=NA,
    annCol=meta[,
		c("Group", "Timepoint")],
    annLegend=FALSE, 
    annColors=ann_colors,
    main="Correlation plot")
```

We can already see interesting patterns emerging from the correlation plot.
First, the cross-correlation amongst samples taking from the control mice is
higher than the cross correlation amongst the rest of the treatments. Second,
the influenza-infected mice midly react until time point 36. Third, the
California strain samples seem closer to control samples than the
lowly-pathogenic Kawasaki samples. Fourth, the Vietnam samples at time point
120 and 168 are the one that are the most different from control samples.