### 2019-05-02 scripts/00_introduction.Rmd
Adding schemas of the workflow + rewording

---
title: "Introduction: a pipeline to analyse time-course gene expression data"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
---

# Abstract

The phenotypic diversity of cells is governed by a complex equilibrium between
their genetic identity and their environmental interactions: Understanding the
dynamics of gene expression is a fundamental question of biology. However,
analysing time-course transcriptomic data raises unique challenging
statistical and computational questions, requiring the development of novel
methods and software.  Using as case study time-course transcriptomics data
from mice exposed to different strains of influenza, ihis workflows provides a
step-by-step tutorial of the methodology used to analyse time-course data:
(1) normalization of the micro-array dataset; (2) differential expression
analysis using functional data analysis; (3) clustering fo time-course data;
(4) interpreting clusters with GO term and KEGG pathway enrichment analysis.

# Introduction

Gene expression studies provide simultaneous quantification of the level of
mRNA from all genes in a sample. High-throughput studies of gene expression
have a long history, starting with microarray technologies in the 1990s
through to single-cell technologies. While many expression studies are
designed to compare the gene expression in distinct groups, there is also a
long history of time-course expression studies, where the the gene expression
is compared across time by measuring mRNA levels from different samples across
time [^1]. Such time course studies can vary from measuring a few distinct time
points, to sampling ten to twenty time points. Many longer time series are
particularly interested in investigating development over time. More recently,
single-cell studies track single cells through their development, and a single
cell is measured at a particular moment in its developmental progression -- a
value that is not know but estimated from the data as its "pseudo-time."

[^1]: Because the collection of the mRNA is often destructive, samples at
different time points are generally from different biological samples;
longitudinal studies, for example tracking the same subject over time, are
certainly possible in certain settings, but not directly considered here.

While there are many methods that have been proposed for discrete aspects of
time course data, the entire workflow for analysis of such data remains
difficult, particularly for long, developmental time series. Most methods
proposed for time course data are concerned with detecting genes that are
changing over time (differential expression analysis), examples being `edge`
[@storey:significance], functional component analysis based models [@wu:more],
time-course permutation tests [@park:statistical], and multiple testing
strategies to combine singe time point differential expression analysis
[@sun:multiple]. However, with long time course datasets, particularly in
developmental systems, a massive number of genes will show *some* change. For
example, in the \xxx system, over \xxx\% of genes are shown to be changing
over time. The task in these settings is often not to detect changes in genes,
but to categorize into biologically interpretable patterns the vast number of
changes discovered.

We present here a workflow for such an analysis that consists of 4 main
parts (Figure \ref{fig:workflow}):

- Quality control and normalization;
- Identification of genes that are differentially expressed;
- Clustering of genes into distinct temporal patterns;
- Biological interpretation of the clusters.

This workflow represents an integration of both novel implementations of
previously established methods and new methodologies for the settings of
developmental time series. It relies on several standard packages for
analysing gene expression data, some specific for time-course data, others
broadly used by the community. We provide the various steps of the workflow as
functions in a R package **FIXME**.

```{r schema, echo=FALSE, fig.width=8, fig.cap="Workflow for analyzing time-course datasets."}
library(knitr)
knitr::include_graphics("images/fig1_schemas.png")
```

# Analysis of the dynamical response of mouse lung tissue to influenza

## Overview

This workflow is illustrated using data from a micro-array time-course
experiment, exposing mice to three different strains of influenza, and
collecting lung tissue during 14 time-points after infection (0, 3, 6, 9, 12,
18, 24, 30, 36, 48, 60 hours, then 3, 5, and 7 days later)
[@shoemaker:ultrasensitive]. The three strains of influenza used in the
study are (1) a low pathogenicity seasonal H1N1 influenza virus
(A/Kawasaki/UTK4/2009 [H1N1]), a mildly pathogenic virus from the 2009
pandemic season (A/California/04/2009 [H1N1]), and a highly pathogenic H5N1
avian influenza virus (A/Vietnam/1203/2004 [H5N1]. Mice were injected with
10^5 PFU of each virus. An adidtional 42 mice were injected with a lower dose
of the Vietnam avian influenza virus (10^3 PFU).

By combining gene expression time-course data with virus growth data, the
authors show that the inflammatory response of lung tissue is gated until a
threshold of the virus concentration is  exceeded in the lung. Once this
threshold is exceeded, a strong inflammatory and cytokin production occurs.
This results provides evidence that the pathology response is non-linearly
regulated by virus concentration.

## References