### 2019-05-03 scripts/04_downstream_analysis.Rmd
Switch bootstrapping proportion

---
title: "Downstream analysis: A pipeline to analyse time-course gene expression data"
output: rmarkdown::html_vignette
---

```{r echo=FALSE, results="hide", message=FALSE}
# Loading dependencies
library(devtools)
library(limma)
library(splines)
library(stats)
library(KEGGprofile)
library(kableExtra)
library(viridis)

library(moanin)
```

```{r echo=FALSE, results="hide"}
# Set options
df = 6

# Preprocessing & filtering
is_count = FALSE
take_log = FALSE    
filter_expression = FALSE

# Differential expression analysis
timecourse_contrasts = c("C-K", "C-M")

# Clustering
filter_genes = TRUE
n_genes_to_keep = 5000
n_clusters = 20
percentage_genes_to_label = 0.5
```

## Downstream analysis of clusters.

Once good clusters are obtained, the next step is to leverage the clustering
to ease interpretation. Classic enrichment analysis step can then be performed
in the gene set defined in each cluster: KEGG pathway enrichment analysis, GO
term enrichment analysis, motif enrichment analysis, etc.

```{r echo=FALSE, results="hide"}
# Set data files and options
data("shoemaker2015")
data = shoemaker2015$data
meta = shoemaker2015$meta

labels = read.table(".results/clustering_labels.txt", sep="\t",
		    check.names=FALSE)
```

First, let us clean up the gene obtained and only select the genes we are
going to use in the enrichment analysis. One can either use the whole set of
genes, only the set of differentially expressed genes in each cluster, or a
subset of genes that fit well to a cluster (based on some criterion).

```{r}
gene_names = row.names(labels)
gene_names = gene_names[!is.na(labels)]
labels = labels[!is.na(labels)]
```

### Finding enriched pathways using `biomaRt` and `KEGGprofile`

Let us first tackle the case of pathway enrichment analysis. We will leverage
the packages `biomaRt` and `KEGGprofile` for this step. `KEGGprofile` is a
package that easily allows to perform pathway enrichment analysis on a set of
genes labeled with the ensembl annotation. We thus need to convert the gene
names into the appropriate format. This is where `biomaRt` comes in handy: it
enables easy conversion from one gene annotation to another. Here, we will use
it to convert the gene names from the Refseq annotation to the ensembl one.

```{r results="hide"}
library(biomaRt)
ensembl = biomaRt::useMart("ensembl")
ensembl = biomaRt::useDataset("mmusculus_gene_ensembl", mart=ensembl)
```

```{r enrichment_analysis, results="hide", message=FALSE}
cluster = 13
genes = gene_names[labels == cluster]

# convert gene names
genes = biomaRt::getBM(attributes=c("ensembl_gene_id", "entrezgene"),
		filters="refseq_mrna", values=genes,
		mart=ensembl)["entrezgene"]
genes = as.vector(unlist(genes))
pathways = KEGGprofile::find_enriched_pathway(
    genes, species="mmu",
    download_latest=TRUE)$stastic

```

```{r echo=FALSE}

pathways = pathways[
    order(pathways["pvalueAdj"]),]

# Select the columns to print
pathways = subset(
    pathways,
    select=c(
	"Pathway_Name", "Percentage",
	"pvalueAdj"))

colnames(pathways) = c("Pathway", "Percentage", "Adj. p-value")
pathways["Percentage"] = pathways["Percentage"] * 100
kable(pathways, "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```

### Finding enriched GO terms

To find GO terms, we use `biomaRt` to find the mapping between GO terms and
gene mapping. The GO enrichment library `topGO` expects the GO term to gene
mapping to be a list where each item is a mapping between a gene name and a GO
term ID vector.

```text
$NM_199153
[1] "GO:0016020" "GO:0016021" "GO:0007186" "GO:0004930" "GO:0007165"
[6] "GO:0050896" "GO:0050909"

$NM_201361
 [1] "GO:0016020" "GO:0016021" "GO:0003674" "GO:0008150" "GO:0005794"
 [6] "GO:0005829" "GO:0005737" "GO:0005856" "GO:0005874" "GO:0005739"
[11] "GO:0005819" "GO:0000922" "GO:0072686"
```

`biomaRt` queries results a matrix with two named columns of gene names and GO
term ID. **FIXME** matrix or data.frame??

```{r go_to_gene_mapping, results="hide", message=FALSE}
# FIXME move this to the package
genes = biomaRt::getBM(attributes=c("go_id", "refseq_mrna"),
	      values=gene_names,
	      filters="refseq_mrna",
	      mart=ensembl)

# Create gene to GO id mapping
gene_id_go_mapping = create_go_term_mapping(genes)
```

Once the gene ID to GO mapping list is created, `moanin` provides a simple
interface to `topGO` to fetch enriched GO terms. Here, we show an example of
running a GO term enrichment on the "Biological process" ontology (BP).

```{r go_term_enrichment, results="hide", message=FALSE}
go_terms_enriched = find_enriched_go_terms(
    scores,
    13,
    gene_id_go_mapping, ontology="BP")
```

```{r echo=FALSE}

go_terms_enriched = go_terms_enriched[
    order(go_terms_enriched[, "resultFisher_padj"]), ]

kable(go_terms_enriched, "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```