### 2019-05-13 scripts/miscellaneous/supp_scores.Rmd

---
title: "Clustering: A pipeline to analyse time-course gene expression data"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
---

```{r echo=FALSE, results="hide", message=FALSE}
# Loading dependencies
library(devtools)
library(limma)
library(splines)
library(stats)
library(moanin)
library(NMF)
source("utils.R")
# library(moanin)
```

```{r echo=FALSE, results="hide"}
# Set options
df = 6

# Preprocessing & filtering
is_count = FALSE
take_log = FALSE    
filter_expression = FALSE

###############################################################################
# Clustering

# Whether to filter genes
filter_genes = TRUE
keep_all_significant = FALSE
n_genes_to_keep = 5000
n_clusters = 20
percentage_genes_to_label = 0.5
```

```{r echo=FALSE, results="hide"}
# Set data files and options
data("shoemaker2015")

data = shoemaker2015$data
meta = shoemaker2015$meta

data = read.table(".results/quantile_normalized.txt")

# Right now, the package doesn't contain the normalized data
de_analysis = read.table(".results/pvalues.txt", sep="\t", check.names=FALSE)

# Extract pvalues and log_fold change
lfc_col_to_keep = colnames(de_analysis)[
    grepl("-lfc_max", colnames(de_analysis))]
lfc_max = de_analysis[, lfc_col_to_keep]

pval_col_to_keep = colnames(de_analysis)[
    grepl("-pval", colnames(de_analysis))]
pvalues = de_analysis[, pval_col_to_keep]

splines_model = moanin::create_splines_model(meta, degrees_of_freedom=6)
```

# Clustering of time-course data

The next step is to perform clustering. First, we reduce the set of genes of
interest to genes that are (1) significant; (2) "highly" differentially
expressed. First, we select the genes that have a log-fold change of at least
2 across at least one experiment for one time point. We thus use the object
`lfc_max`.

XXX

```{r filter_genes}
if(filter_genes){
    # First, filter out any genes that doesn't have a log fold change of at
    # least 2 at at least one of the time points.

    genes_to_keep = row.names(lfc_max[rowSums(lfc_max > 2) > 0, ])
    pvalues = pvalues[genes_to_keep, ]

    if(length(genes_to_keep) > n_genes_to_keep){

	# Then rank by fisher's p-value and take max the number of genes of
	# interest
	# Filter out q-values for the pvalues table
	fishers_pval = moanin:::fisher_method(pvalues)
	fishers_qval = stats::p.adjust(fishers_pval)
        if(keep_all_significant){
	   genes_to_keep = names(fishers_qval[fishers_qval < 0.05])
	}else{
	    genes_to_keep = names(sort(fishers_pval)[1:n_genes_to_keep])
	}
    }

    y = as.matrix(data[genes_to_keep, ])
}else{
    y = as.matrix(data)
}
```

After filtering, we consider `r dim(y)[1]` genes for the clustering.

```{r clustering}

#############################################################################
# Fit the splines

kmeans_clusters = moanin:::splines_kmeans(
    y, splines_model, n_clusters=20,
    random_seed=42,
    n_init=20)

# XXX is there a bug here?
all_scores = data.frame(row.names=row.names(data))
for(k in 1:20){
     scores = moanin:::score_genes_centroid(data, kmeans_clusters$centroids[k,])
     all_scores[paste0("cluster_", k)] = scores
 }

# Only assign labels to X% of the genes
max_score = quantile(moanin:::row_min(all_scores), c(percentage_genes_to_label))
genes_to_not_consider = moanin:::row_min(all_scores) >= max_score
labels = moanin:::row_argmin(all_scores)
labels[genes_to_not_consider] = NA
```

```{r, fig.width=4, fig.height=4}
hist(all_scores[, 1], col="black")
hist(all_scores[, 2], col="black")

```

```{r}
table(kmeans_clusters$clusters)
```

```{r}
table(labels)
```